Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy

### 2021-04-17 manuscript/manuscript.Rmd
Make manuscript work with targets

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "October 2, 2020"
published: Working paper.
git-repo: https://github.com/andrewheiss/who-cares-about-crackdown
abstract: >-
  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals’ preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
reference-section-title: References
toc: false
title-page: true
word-count: "7,562"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
thanks: >-
  The data that support the findings of this study, as well as the code to reproduce the findings, are openly available at <https://github.com/andrewheiss/who-cares-about-crackdown> and <https://osf.io/sm5ew/>. This experiment received approval by the human subjects research committees at Christopher Newport University (1436622-1), Brigham Young University (E19135), and Georgia State University (H19644). As part of a larger project, we preregistered hypotheses and our research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/hsbyd/>. This study covers the third research question of our preregistered design.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Marc Dotson, PhD, Assistant Professor of Marketing, Marriott School of Business, Brigham Young University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
mainfont: Lora
sansfont: IBM Plex Sans
policy-implications: >-
  In the wake of the global crackdown on civil society, international NGOs (INGOs) need to maintain access to multiple sources of funding, including individual donors.

@@

  Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences.

  We find that potential donors with low social trust tend to be wary of negative host-country relationships and prefer to donate to INGOs that are friendly with their host governments, while potential donors with high social trust are more willing to stick with legally besieged INGOs.

  Donors that are the most trusting and involved may remain supportive when charities face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors.
---

@@

@@

```{r child = "output/tables/tbl-organization-attributes-full.md"}
```

@@

The hypothetical organizations are based on random combinations of features, including one of four issue areas, the presence or absence of two organizational practices, one of three funding sources, and one of three relationships with host governments (see @tbl:organization-attributes-full; see also Q3.1 in the survey text in the appendix). Respondents saw twelve sets of four hypothetical organizations out of 288 (4 × 4 × 2 × 3 × 3) possible combinations, meaning no single respondent was offered every combination. While not every combination necessarily reflected reality—for instance, a respondent might have been asked to compare a fictional Amnesty International that focuses on environmental issues and that is funded by private donations with a Greenpeace that focuses on human rights and that is funded by government grants—it is not important that these features align with true organizational characteristics, since the goal of these repeated hypothetical questions is to identify which attributes are the most salient for donors.

@@

```{r child = "output/tables/tbl-sample-details.md"}
```

@@

@tbl:sample-details provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer and donate to charity at least once a year and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and relationships with host governments when deciding to donate. Disentangling the effect of every individual level-characteristic measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. To simplify the possible combinations of individual characteristics, we generate 64 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see @tbl:persona-attributes). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see @tbl:organization-attributes).

@@

```{r child = "output/tables/tbl-persona-attributes.md"}
```

```{r child = "output/tables/tbl-organization-attributes.md"}
```

@@

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in @tbl:persona-attributes are held at their means or modal values from the survey sample. @tbl:sim-output provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation.

@@

```{r child = "output/tables/tbl-sim-output.md"}
```

@@

## Demographics

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The four panels of @fig:income show the average predicted market share for different organizational issue areas, funding sources, and relationships with host governments across different levels of persona income, with two different cutpoints for high and low income. To ease with modeling and interpretation, we split high- and low-income personas based on the 2017 U.S. median income (\$61,372), showing personas with incomes higher and lower than the national median. Because there is wide variation in donor incomes in the U.S., we also show the results from a simulation model with a more extreme split in income, holding annual persona income at \$50,000 and \$100,000+. The difference in the division of high and low income makes no difference in the market shares across issue areas and funding sources, but there are substantial income-based differences across different relationships with host governments.

![Average predicted donation market shares across all personas, segmented by persona income across organizational issue area, relationship with host government, and funding sources](output/figures/income-all.pdf){#fig:income}

@@

![Average predicted donation market shares across all personas, segmented by persona education and religiosity across organizational issue area, relationship with host government, and funding sources](output/figures/education-religion-all.pdf){#fig:edu-religion}

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns (as seen in @fig:edu-religion). Human rights issues attract a slightly larger market share of personas who are college graduates or who attend religious services at least monthly, while refugee issues are more likely to attract less religious personas, but these differences are minor. Preferences are consistent for different sources of funding as well—personas generally prefer organizations funded mostly by government grants. Persona preferences across different organizational relationships with host governments are also consistent across education and religiosity, with most personas preferring to donate to NGOs that enjoy friendly relations abroad.

## Public affairs knowledge, experience, and social trust

@@

![Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different NGO issue areas](output/figures/issue-social.pdf){#fig:issue-social}

@fig:issue-social and @fig:relationship-social show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In @fig:issue-social, those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organizations. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue.

>>>>

---
title: "Who Cares About Crackdowns? Exploring the Role of Trust in Individual Philanthropy"
short-title: "Who Cares About Crackdowns?"
author:
- name: Suparna Chaudhry
  affiliation: Lewis & Clark College
  url: http://www.suparnachaudhry.com/
  email: schaudhry@lclark.edu
- name: Marc Dotson
  affiliation: Brigham Young University
  url: https://marriottschool.byu.edu/directory/details?id=50683
  email: marc_dotson@byu.edu
- name: Andrew Heiss
  affiliation: Georgia State University
  url: https://www.andrewheiss.com/
  email: aheiss@gsu.edu
date: "October 2, 2020"
published: Working paper.
git-repo: https://github.com/andrewheiss/who-cares-about-crackdown
abstract: >-
  The phenomenon of closing civic space has adversely impacted the operations of INGOs. It has not just resulted in net losses of income to INGOs, but also how and where funds are allocated. We argue that individual private donors can be important in sustaining the operations of INGOs working in repressive contexts. Individual donors do not use the same performance-based metrics as official aid donors, and trust has shown to be an important component of support for nonprofits working towards difficult goals. How does trust in charitable organizations influence individuals’ preferences to donate, especially when these groups face crackdown? Using a simulated market for philanthropic donations towards INGOs, based on data from a nationally representative sample of individuals in the United States who regularly donate to charity, we find that trust in charitable organizations matters substantially in shaping donor preferences. Donor profiles with high levels of social trust are likely to donate to INGOs with friendly relationships with host governments. This support holds steady if INGOs face criticism or crackdown. In contrast, donor profiles with lower levels of social trust prefer to donate to organizations that do not face criticism or crackdown abroad. The global crackdown on NGOs may thus possibly sour NGOs' least trusting individual donors. Our findings have practical implications for INGOs raising funds from individuals amid closing civic space.
toc: false
reference-section-title: References
bibliography: bibliography.bib
title-page: true
word-count: "7,562"
keywords: international NGOs, donor behavior, philanthropy, human rights, repression, international giving, donor choice
thanks: >-
  The data that support the findings of this study, as well as the code to reproduce the findings, are openly available at <https://github.com/andrewheiss/who-cares-about-crackdown> and <https://osf.io/sm5ew/>. This experiment received approval by the human subjects research committees at Christopher Newport University (1436622-1), Brigham Young University (E19135), and Georgia State University (H19644). As part of a larger project, we preregistered hypotheses and our research design at the Open Science Framework, and our preregistration protocol is available at <https://osf.io/hsbyd/>. This study covers the third research question of our preregistered design.
author-note: >-
  Suparna Chaudhry, PhD, Assistant Professor of International Affairs, Department of International Affairs, Lewis \& Clark College; Marc Dotson, PhD, Assistant Professor of Marketing, Marriott School of Business, Brigham Young University; Andrew Heiss, PhD, Assistant Professor of Public Management, Andrew Young School of Policy Studies, Georgia State University. 
correspondence: >-
  Correspondence concerning this article should be addressed to Andrew Heiss, PhD, Department of Public Management and Policy, Andrew Young School of Policy Studies, Georgia State University, 14 Marietta Street NW, Suite 357, Atlanta, GA 30303. Contact: aheiss@gsu.edu
mainfont: Lora
sansfont: IBM Plex Sans
policy-implications: >-
  In the wake of the global crackdown on civil society, international NGOs (INGOs) need to maintain access to multiple sources of funding, including individual donors. 

@@

  Trust in political institutions and in charitable organizations matters substantially in shaping donor preferences.
  We find that potential donors with low social trust tend to be wary of negative host-country relationships and prefer to donate to INGOs that are friendly with their host governments, while potential donors with high social trust are more willing to stick with legally besieged INGOs.
  Donors that are the most trusting and involved may remain supportive when charities face government criticism and crackdown. The global crackdown on NGOs may also possibly sour NGOs' least trusting individual donors.
editor_options: 
  chunk_output_type: console
---

@@

```{r setup, include=FALSE}
# Figure out output format
is_docx <- knitr::pandoc_to("docx") | knitr::pandoc_to("odt")
is_latex <- knitr::pandoc_to("latex")
is_html <- knitr::pandoc_to("html")

# Word-specific things
table_format <- ifelse(is_docx, "huxtable", "kableExtra")  # Huxtable tables
conditional_dpi <- ifelse(is_docx, 300, 300)  # Higher DPI
conditional_align <- ifelse(is_docx, "default", "center")  # Word doesn't support align

# Knitr options
knitr::opts_chunk$set(
  echo = FALSE, warning = FALSE, message = FALSE,
  tidy.opts = list(width.cutoff = 120),  # Code width
  fig.retina = 3, dpi = conditional_dpi,
  fig.width = 7, fig.asp = 0.618,
  fig.align = conditional_align, out.width = "100%",
  fig.path = "output/figures/",
  cache.path = "output/_cache/",
  fig.process = function(x) {  # Remove "-1" from figure names
    x2 = sub('-\\d+([.][a-z]+)$', '\\1', x)
    if (file.rename(x, x2)) x2 else x
  }
)

# R options
options(
  width = 90,  # Output width
  dplyr.summarise.inform = FALSE,  # Turn off dplyr's summarize() auto messages
  knitr.kable.NA = "",  # Make NAs blank in kables
  kableExtra.latex.load_packages = FALSE,  # Don't add LaTeX preamble stuff
  modelsummary_default = table_format  # Set modelsummary backend
)
```

```{r libraries-data, include=FALSE}
library(tidyverse)
library(targets)
library(kableExtra)
library(patchwork)
library(scales)
library(janitor)
library(glue)
library(here)

withr::with_dir(here::here(), {
  source(tar_read(plot_funs))

  tar_load(personas)
  tar_load(orgs)
  tar_load(participant_summary)
  tar_load(sim_final)
  tar_load(sim_excel_final)
})
```

@@

```{r organization-attributes-full}
orgs$org_attributes %>% 
  select(Organization, `Issue area`, `Organizational practices`, 
         `Funding sources`, `Relationship with government`) %>% 
  kbl(align = "lllll",
      caption = "Organization attributes varied in the experiment. Each participant saw repeated combinations of attributes, with one attribute randomly selected from each column.") %>% 
  kable_styling()
```

@@

The hypothetical organizations are based on random combinations of features, including one of four issue areas, the presence or absence of two organizational practices, one of three funding sources, and one of three relationships with host governments (see Table \@ref(tab:organization-attributes-full); see also Q3.1 in the survey text in the appendix). Respondents saw twelve sets of four hypothetical organizations out of 288 (4 × 4 × 2 × 3 × 3) possible combinations, meaning no single respondent was offered every combination. While not every combination necessarily reflected reality—for instance, a respondent might have been asked to compare a fictional Amnesty International that focuses on environmental issues and that is funded by private donations with a Greenpeace that focuses on human rights and that is funded by government grants—it is not important that these features align with true organizational characteristics, since the goal of these repeated hypothetical questions is to identify which attributes are the most salient for donors.

@@

```{r sample-details}
participant_summary %>% 
  select(Question = clean_name,
         Response = level,
         N = count,
         `%` = nice_prop) %>% 
  kbl(align = "lllcc",
      caption = "Summary of individual respondent characteristics") %>% 
  pack_rows(index = table(fct_inorder(participant_summary$category))) %>%
  collapse_rows(columns = 1, valign = "top") %>% 
  kable_styling()
```

@@

Table \@ref(tab:sample-details) provides a summary of the individual characteristics of survey respondents. Because we limited the sample to people who regularly make charitable contributions, most respondents have a favorable opinion of nonprofits and charities, with more than half reporting a 5 or greater level of trust in charities (on a 7-point scale). Moreover, nearly 90% indicated that it is important to trust charities. Additionally, the majority of respondents volunteer and donate to charity at least once a year and give between \$1–\$999 annually. Most rarely attend church, but feel that religion is important. Respondents are also civically engaged (three quarters voted in the last election) and follow national news closely (70% follow the news daily). Respondents are fairly balanced ideologically, with the majority reporting a 3, 4, or 5 on a 7-point scale of liberal–conservative political ideology.

@@

In this paper, we look at which kinds of donors are more or less likely to consider an NGO's issue area, funding sources, and relationships with host governments when deciding to donate. Disentangling the effect of every individual level-characteristic measured in the survey on every combination of organizational characteristics presented in the conjoint experiment is unwieldy and near impossible. Given the organizational attributes presented in our experiment, and the possible individual characteristics that respondents could select, there are billions of possible combinations and iterations of organizational and individual attributes—many of which rarely occur in the actual population of potential donors. To more fully explore the relationships between different organizational- and individual-level attributes, we turn to simulation methods that allow us to examine the impact of multiple covariates simultaneously. To simplify the possible combinations of individual characteristics, we generate 64 distinct personas with attributes that vary along different dimensions of demographics, politics and public affairs experience, and social attitudes (see Table \@ref(tab:persona-attributes)). We also generate 24 simulated organizations that vary by issue area, relationship with host government, and source of funding (see Table \@ref(tab:organization-attributes)).

@@

```{r persona-attributes}
personas$persona_attributes %>% 
  kbl(align = "lll",
      caption = "Individual attributes varied in the simulation, resulting in 64 persona profiles") %>% 
  kable_styling()
```

```{r organization-attributes}
orgs$org_attributes %>% 
  select(`Issue area`, `Relationship with government`, `Funding`) %>% 
  kbl(align = "lll",
      caption = "Organization attributes varied in the simulation, resulting in 24 hypothetical organizations") %>% 
  kable_styling()
```

@@

We then combine these different scenarios of donor and organization attributes with the coefficients from a multinomial multilevel regression model to calculate the probability that each simulated persona has of giving to specific combinations of organizational attributes. Any individual characteristics that we do not explicitly vary in Table \@ref(tab:persona-attributes) are held at their means or modal values from the survey sample. Table \@ref(tab:sim-output) provides an example of the simulation output, showing how two different personas are predicted to allocate donations to organizations with various characteristics. The proportion of donations from each persona sums to 100%, as every persona is guaranteed to make a donation.

@@

```{r sim-output}
example_personas <- c("persona2", "persona63")

example_persona_results <- sim_final %>% 
  filter(persona_id %in% example_personas) %>% 
  mutate(org_funding = str_to_sentence(str_remove(org_funding, "Mostly funded by "))) %>% 
  mutate(org_clean = glue("{organization}: {org_issue}, {org_funding}, {org_relationship}")) %>% 
  mutate(persona_desc = recode(
    persona_id,
    "persona2" = "Lower income high school graduate who rarely attends religious services; liberal who reads and travels; doesn't trust or donate",
    "persona63" = "Higher income college graduate who attends religious services; conservative who doesn't read or travel; trusts and donates")
  ) %>%
  mutate(persona_clean = glue("{persona}: {persona_desc}")) %>% 
  select(persona_clean, share, org_clean) %>% 
  pivot_wider(names_from = "persona_clean", values_from = "share") %>% 
  adorn_totals(where = "row", name = "Total")

example_persona_results_small <- example_persona_results %>% 
  slice(c(1, 2, 3, 7, 8, 9, 16, 17, 25)) %>% 
  mutate(across(where(is.numeric), ~ percent_format(accuracy = 0.1)(.))) %>% 
  add_row(org_clean = "…", .after = 3) %>% 
  add_row(org_clean = "…", .after = 7) %>% 
  add_row(org_clean = "…", .after = 10) %>% 
  mutate(across(everything(), ~replace_na(., "…"))) %>% 
  rename(Organization = org_clean)

example_persona_results_small %>% 
  mutate(Organization = text_spec(Organization, bold = Organization == "Total")) %>% 
  kbl(align = "lcc",
      caption = "Sample simulation output",
      escape = FALSE) %>% 
  kable_styling()
```

@@

## Patterns across demographics

We begin by looking at differences in donation preferences across a range of basic demographic characteristics. The four panels of Figure \@ref(fig:income) show the average predicted market share for different organizational issue areas, funding sources, and relationships with host governments across different levels of persona income, with two different cutpoints for high and low income. To ease with modeling and interpretation, we split high- and low-income personas based on the 2017 U.S. median income (\$61,372), showing personas with incomes higher and lower than the national median. Because there is wide variation in donor incomes in the U.S., we also show the results from a simulation model with a more extreme split in income, holding annual persona income at \$50,000 and \$100,000+. The difference in the division of high and low income makes no difference in the market shares across issue areas and funding sources, but there are substantial income-based differences across different relationships with host governments.

```{r income, fig.width=6, fig.height=4.5, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona income across organizational issue area, relationship with host government, and funding sources"}
plot_income_issue <- sim_final %>% 
  group_by(org_issue, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Issue area") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)),
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002)),
                     breaks = seq(0, 0.06, 0.02)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], guide = FALSE) +
  coord_cartesian(xlim = c(0, 0.06)) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_income_funding <- sim_final %>% 
  group_by(org_funding, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Funding sources") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_funding, 10)), 
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_income_relationship <- sim_final %>% 
  mutate(persona_income = fct_recode(persona_income,
                                     "< $61,372/year" = "Lower income",
                                     "> $61,372/year" = "Higher income")) %>% 
  group_by(org_relationship, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5, 
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) +
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.003))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], 
                     guide = guide_legend(reverse = TRUE, nrow = 1,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0))) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 50))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

plot_income_relationship_extreme <- sim_excel_final %>% 
  mutate(persona_income = fct_recode(persona_income,
                                     "$50,000/year" = "Lower income",
                                     "$100,000/year" = "Higher income"),
         org_relationship = fct_recode(org_relationship,
                                       "Under crackdown" = "Crackdown")) %>% 
  group_by(org_relationship, persona_income) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = fct_rev(persona_income))) +
  geom_pointrange(size = 0.75, fatten = 1.5, linetype = "21", pch = 2,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[1]], 
                     guide = guide_legend(reverse = TRUE, nrow = 1,
                                          override.aes = list(size = 0.25, 
                                                              linetype = 0))) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 50))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

plot_income <- ((plot_income_issue + labs(x = NULL)) + 
  (plot_income_funding + labs(x = NULL))) / 
  (plot_income_relationship + plot_income_relationship_extreme)

plot_income
```

@@

```{r edu-religion, fig.width=6, fig.height=4.5, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona education and religiosity across organizational issue area, relationship with host government, and funding sources"}
# Education plots
plot_education_issue <- sim_final %>% 
  group_by(org_issue, persona_education) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Issue area") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)), 
             x = avg_share, color = fct_rev(persona_education))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[2]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_education_funding <- sim_final %>% 
  group_by(org_funding, persona_education) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Funding sources") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_funding, 10)), 
             x = avg_share, color = fct_rev(persona_education))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[2]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_education_relationship <- sim_final %>% 
  group_by(org_relationship, persona_education) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)),
             x = avg_share, color = fct_rev(persona_education))) +
  geom_pointrange(size = 0.75, fatten = 1.5,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[2]], 
                     guide = guide_legend(reverse = TRUE, ncol = 1,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0)),
                     labels = label_wrap(15)) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 20))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

# Religion plots
plot_religion_issue <- sim_final %>% 
  group_by(org_issue, persona_religion) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Issue area") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)), 
             x = avg_share, color = fct_rev(persona_religion))) +
  geom_pointrange(size = 0.75, fatten = 1.5, pch = 17,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[4]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_religion_funding <- sim_final %>% 
  group_by(org_funding, persona_religion) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Funding sources") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_funding, 10)), 
             x = avg_share, color = fct_rev(persona_religion))) +
  geom_pointrange(size = 0.75, fatten = 1.5, pch = 17,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.06)) +
  scale_color_manual(values = clrs_ngo_pairs[[4]], guide = FALSE) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(facet)) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank())

plot_religion_relationship <- sim_final %>% 
  group_by(org_relationship, persona_religion) %>% 
  summarize(avg_share = mean(share)) %>% 
  mutate(facet = "Relationship with host government") %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_relationship, 10)), 
             x = avg_share, color = fct_rev(persona_religion))) +
  geom_pointrange(size = 0.75, fatten = 1.5, pch = 17,
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  coord_cartesian(xlim = c(0, 0.1)) +
  scale_color_manual(values = clrs_ngo_pairs[[4]], 
                     guide = guide_legend(reverse = TRUE, ncol = 1,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0)),
                     labels = label_wrap(20)) +
  labs(x = "Average donation share", y = NULL, color = NULL) +
  facet_wrap(vars(str_wrap(facet, 20))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))

plot_education_religion <- (plot_education_issue + labs(x = NULL)) + 
  (plot_education_funding + labs(x = NULL)) + 
  (plot_education_relationship + labs(x = NULL)) + 
  guide_area() +
  plot_religion_issue + 
  plot_religion_funding + 
  plot_religion_relationship +
  plot_layout(guides = "collect", ncol = 4)

plot_education_religion
```

While education and religiosity are both important drivers of individual philanthropy, neither are associated with any sizable differences in persona donation patterns (as seen in Figure \@ref(fig:edu-religion)). Human rights issues attract a slightly larger market share of personas who are college graduates or who attend religious services at least monthly, while refugee issues are more likely to attract less religious personas, but these differences are minor. Preferences are consistent for different sources of funding as well—personas generally prefer organizations funded mostly by government grants. Persona preferences across different organizational relationships with host governments are also consistent across education and religiosity, with most personas preferring to donate to NGOs that enjoy friendly relations abroad.

## Patterns across public affairs knowledge, experience, and social trust

@@

```{r issue-social, fig.width=6, fig.height=4, fig.asp=NULL, out.width="100%", fig.cap="Average predicted donation market shares across all personas, segmented by persona public affairs knowledge, political ideology, and social trust across different NGO issue areas"}
sim_final %>% 
  group_by(org_issue, persona_trust, persona_ideology, persona_experience) %>% 
  summarize(avg_share = mean(share)) %>% 
  ggplot(aes(y = fct_rev(str_wrap_factor(org_issue, 15)), 
             x = avg_share, color = persona_trust)) +
  geom_pointrange(size = 0.75, fatten = 1.5, 
                  aes(xmin = 0, xmax = ..x..), position = position_dodge(width = 0.5)) + 
  scale_x_continuous(labels = percent_format(accuracy = 1), expand = expansion(add = c(0, 0.002))) +
  scale_color_manual(values = clrs_ngo_pairs[[3]], 
                     guide = guide_legend(reverse = TRUE,
                                          override.aes = list(size = 0.25,
                                                              linetype = 0))) +
  labs(x = "Average share of donations", y = NULL, color = NULL) +
  facet_grid(rows = vars(persona_ideology), cols = vars(str_wrap(persona_experience, 100))) +
  theme_ngo() +
  theme(panel.grid.major.y = element_blank(),
        legend.key.width = unit(0.5, "lines"))
```

Figures \@ref(fig:issue-social) and \@ref(fig:relationship-social) show the results of the market simulation across these three dimensions of persona characteristics, varying public affairs knowledge, political ideology, and social trust. Personas respond differently to organizational issue areas. In Figure \@ref(fig:issue-social), those who follow the news and travel abroad are more likely to prefer donating to human rights organizations than their less knowledgeable and less traveled counterparts. Personas who do not follow the news or travel show a greater preference for emergency response and refugee relief organizations. There are no strong ideological effects across issue areas. Both conservative and liberal personas tend to follow similar patterns of preference across NGO issue.
